Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-01 mu/article/article.tex
Making last corrections.

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage{hyperref}   % use for hypertext links, including those to external 
                        % documents and URLs 
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
%\usepackage[inline]{showlabels}
\input{/Users/bms/util/definitions}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%                         %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%% I N T R O D U C T I O N %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%                         %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state 
media plays the determining role, adding a new degree of freedom to
conventional charge manipulation.\cite{wolfSC04,fabianAPS07}
% 
At present, there is an increasing interest in attaining the same level of
control over the transport of spin at micro or nano scales, as it has been done for
the flow of charge in typical electronic devices.\cite{awschalomNP2007} Some
semiconductor spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC).
% 
One of the difficulties to achieve the development of spin current and PSC
semiconductor devices is the fact that the spin relaxation time in a
semiconducting media is short, disabling the spin transport, and then resulting
in a non-observable spin current.\cite{murakamiSc03}
% 
In PSCs, there is no net
motion of charge; spin-up electrons move in a given direction, while spin-down
electrons travel in the opposite one. 
This effect can result from one photon absorption of linearly polarized light
by a semiconductor, with filled valence bands and empty conduction bands,
illuminated by light with photon energy larger than the energy gap.
This phenomenon can result from spin
injection,\cite{malPRB03} Hall Effects,\cite{sinovaPRB04} interference of two
optical beams,\cite{bhatPRL00, najmaiePRB03} or one photon absorption of
linearly polarized light\cite{bhatPRL05} and has been observed in gallium
arsenide (GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

Graphene, an allotrope of carbon with hexagonal 2D lattice structure, presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08}, being then a
perfect platform to be used in two-dimension electronic systems; however, most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper, we present two 50\% hydrogenated graphene noncentrosymmetric
structures, both presenting a discernible band gap: the Up structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only on the upper side of the structure, while the Alt structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/upstruc2}
    \\
    \includegraphics[width=\tama]{figures/upstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Up
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms,
      labeled 1 (2).
}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/altstruc2}
    \\
    \includegraphics[width=\tama]{figures/altstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Alt
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms,
      labeled 1 (2).
}
    \label{fig:alt-struc}
\end{figure}

Using those structures, we address a theoretical study of the 
spin velocity
injection (SVI) 
by one-photon absorption of linearly polarized light.
Since we have 2D structures, we do the analysis for two cases. The first is by
fixing the spin of the electrons along the $z$ Cartesian direction, with the
velocity directed on the surface of the structure on the $xy$ plane. The second
is by fixing the spin velocity in the $x$ or $y$ direction and the spin directed
to $xyz$.
% 
The SVI is an optical effect that quantifies the velocity at which a PSC moves
along the Cartesian direction, $\mathrm{a}$, with the spin of electron polarized
along the Cartesian direction $\mathrm{b}$. One photon absorption of linearly
polarized light can promote an even distribution of electrons in $\mathbf{k}$ space,
regardless of the symmetry of the material, resulting in a 
null electrical current.\cite{bhatPRL05}
Then, the electrons excited to the conduction bands at opposite $\mathbf{k}$
points will result in opposite spin polarizations producing no net spin
injection.\cite{bhatPRL05} If the crystalline structure of the material is
noncentrosymmetric, the spin polarization injected at a given $\mathbf{k}$ point
not necessarily 
 vanishes,\cite{alvaradoPRL85, schmiedeskampPRL88} 
and then, since the velocities of electrons at opposite $\mathbf{k}$ points are
opposite, a PSC will be produced. Since the structures presented
here are noncentrosymmetric, they are good candidates in which this effect can
be induced.

@@

semiconductors. To derive above we used

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of
the electron velocity on the plane of the structure 
with the spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\gs^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
to the incoming polarization, 
respectively. The subscript $\gs^\rmb$ denotes the spin along $\rmb$.

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H & -0.61516 & -1.77416 &  0.73196 \\
1 & H &  0.61518 &  0.35514 &  0.73175 \\
2 & C & -0.61516 & -1.77264 & -0.49138 \\
2 & C & -0.61516 & -0.35600 & -0.72316 \\
2 & C &  0.61516 &  0.35763 & -0.49087 \\
\hline
\end{tabular}

\caption{Unit cell of Up structure. Layer division, atom types and
positions for the Up structure. The structure unit cell was divided in
two layers corresponding to hydrogen and carbon atoms.The corresponding layer
atom position is depicted in Fig. \ref{fig:up-struc} with the corresponding
number of layer.}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H &  -0.61516 &  -1.42140 & \ 1.47237 \\
2 & C &  -0.61516 &  -1.73300 & \ 0.39631 \\
3 & C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
4 & C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
5 & C &  -0.61516 &  -0.37396 &  -0.39632 \\
6 & H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}

@@

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ vs. $\hbar\go$ for the directions $\rma\rmb$ and the angle $\ga$
that maximizes the signal, for the Up  and Alt structures and for CdSe and
GaAs, which are bulk systems shown for comparison.
% 
As expected from the delta function of Eq.~\eqref{eq:mu},
$\calv^{\rma\rmb}(\go,\ga)$ rises right at the corresponding energy gap of each
system. For the 2D structures, the spectrum covers two narrow energy regions
with large values of response, while for bulk systems, the spectra covers a
rather wide energy range, but with a much smaller response.
% 
For the Up structure, $\rma\rmb=yz$ and $\ga=35^\circ$ maximize the response,
which means that an incoming beam of light with its electric field polarized at
35$^\circ$ from the $x$ direction will induce electrons to move along $y$
(parallel to the surface), with their spin polarized along $z$ (perpendicular to
the surface), with the following speeds:
% 
right at the energy onset, $\calv^{yz}(\go,\ga)=668$\,Km/s, a speed that remains
almost constant for 65 meV, and then goes to zero and a second region is found
above 1.946 eV, where there are two extreme values of the speed,
$\calv^{yz}(\go,\ga)=266.3$ Km/s at $\hbar\go=1.954$\,eV, and
$\calv^{\rma\rmb}(\go,\ga)=-241.4$ Km/s at $\hbar\go=1.958$\,eV; a positive
(negative) $\calv^{\rma\rmb}(\go,\ga)$ means that the electrons move parallel
(antiparallel) to the electric field.
Likewise, for the Alt structure, we find that also $\rma\rmb=yz$ and
$\ga=150^\circ$ maximizes the response, where two extreme values of
$\calv^{\rmy\rmz}$ are found, one at  $\hbar\go=0.720$\,eV of
$\calv^{\rmy\rmz}=-711.9$ Km/s, and the other at $\hbar\go=0.911$\,eV of
$\calv^{\rmy\rmz}=-330.6$ Km/s.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{upplots/up-vx-vy-w2}
\caption{For the Up structure we show $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and $\varphi_{\mathrm{a}}
(\omega,\alpha)$, (right scale, blue line), vs. $\hbar\go$, for
$\ga=35^\circ$, and $\rma=x,y$. }
\label{fig:up-vx-vy-w2}
\end{figure}
In Fig. \ref{fig:up-vx-vy-w2}, we plot $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ vs. $\hbar\go$, in the range where there are two local maxima
at $\hbar\go=1.954$\,eV and $\hbar\go=1.957$\,eV.
% 
The former is the largest of the two, with
$\mathcal{V}_{\mathrm{x}} (\omega,\alpha)=61.2$\,Km/s,
$\theta_{\mathrm{x}} (\omega,\alpha)=48.3^{\circ}$, and 
$\varphi_{\mathrm{x}} (\omega,\alpha)=54.3^{\circ}$,
for the electron moving along $x$, and
$\mathcal{V}_{\mathrm{y}} (\omega,\alpha)=293.2$\,Km/s,
$\theta_{\mathrm{y}} (\omega,\alpha)=49.8^{\circ}$, and 
$\varphi_{\mathrm{y}} (\omega,\alpha)=51.9^{\circ}$
for the electron moving along $y$.
% 
For the $\hbar\go=1.957$\,eV peak we obtain 
$\theta_{\mathrm{x}} (\omega,\alpha) = 129.8^{\circ}$, and 
$\varphi_{\mathrm{x}} (\omega,\alpha) = 231.7^{\circ}$, with 
$\mathcal{V}^{\mathrm{x}} (\omega,\alpha) = 54.6$\,Km/s, and
$\theta_{\mathrm{y}}(\omega,\alpha) =129.3$, and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 230.7$, with 
$\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 263.7$\,Km/s. 
Checar que esto sea correcto para ambos maximos!!
We remark that these angles are almost
constant for all the energy values across the peak of these two local maxima, for
which the spin is directed upward in the first Cartesian quadrant of the $xy$
plane when it moves along either $x$ or $y$ directions.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{altplots/alt-vx-vy-w1}
\caption{For the Alt structure we show
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line),
$\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and
$\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue line), vs.
$\hbar\go$, for $\ga=150^\circ$, and $\rma=x,y$. }
\label{fig:alt-vx-vy-w1}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{altplots/alt-vx-vy-w2}
\caption{For the Alt structure we show $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and $\varphi_{\mathrm{a}}
(\omega,\alpha)$, (right scale, blue line), vs. $\hbar\go$, for
$\ga=150^\circ$, and $\rma=x,y$. }
\label{fig:alt-vx-vy-w2}
\end{figure}

>>>>

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage{hyperref}   % use for hypertext links, including those to external 
                        % documents and URLs 
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
%\usepackage[inline]{showlabels}
% \input{/Users/bms/util/definitions}
\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

@@

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state media plays the
determining role, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} 
At present, there is an increasing interest in attaining the same level of
control over the transport of spin at micro or nano scales, as it has been done
for the flow of charge in typical electronic devices.\cite{awschalomNP2007} Some
semiconductor spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC).
One of the difficulties to achieve the development of spin current and PSC
semiconductor devices is the fact that the spin relaxation time in a
semiconducting media is short, disabling the spin transport, and then resulting
in a non-observable spin current.\cite{murakamiSc03}
In PSCs, there is no net motion of charge; spin-up electrons move in a given
direction, while spin-down electrons travel in the opposite one. This effect can
result from one photon absorption of linearly polarized light by a
semiconductor, with filled valence bands and empty conduction bands, illuminated
by light with photon energy larger than the energy gap. This phenomenon can
result from spin injection,\cite{malPRB03} Hall Effects,\cite{sinovaPRB04}
interference of two optical beams,\cite{bhatPRL00, najmaiePRB03} or one photon
absorption of linearly polarized light\cite{bhatPRL05} and has been observed in
gallium arsenide (GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium
arsenide (AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

Graphene, an allotrope of carbon with hexagonal 2D lattice structure, presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08}, being then a
perfect platform to be used in two-dimension electronic systems; however, most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
In this paper, we present two 50\% hydrogenated graphene noncentrosymmetric
structures, both presenting a discernible band gap: the Up structure, shown in
Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer only on
the upper side of the structure, while the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon slab.\cite{zapataPSB2016}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/upstruc2}
    \\
    \includegraphics[width=\tama]{figures/upstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Up
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms. 
}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/altstruc2}
    \\
    \includegraphics[width=\tama]{figures/altstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Alt
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms.
}
    \label{fig:alt-struc}
\end{figure}

Using those structures, we address a theoretical study of the spin velocity
injection (SVI) by one-photon absorption of linearly polarized light. Since we
have 2D structures, we do the analysis for two cases. The first is by fixing the
spin of the electrons along the $z$ Cartesian direction, with the velocity
directed onto the surface of the structure on the $xy$ plane. The second is by
fixing the spin velocity in the $x$ or $y$ direction and the spin directed to
$xyz$.
The SVI is an optical effect that quantifies the velocity at which a PSC moves
along the Cartesian direction, $\mathrm{a}$, with the spin of electron polarized
along the Cartesian direction $\mathrm{b}$. One photon absorption of linearly
polarized light can promote an even distribution of electrons in $\mathbf{k}$
space, regardless of the symmetry of the material, resulting in a null
electrical current.\cite{bhatPRL05} Then, the electrons excited to the
conduction bands at opposite $\mathbf{k}$ points will result in opposite spin
polarizations producing no net spin injection.\cite{bhatPRL05} If the
crystalline structure of the material is noncentrosymmetric, the spin
polarization injected at a given $\mathbf{k}$ point not necessarily
vanishes,\cite{alvaradoPRL85, schmiedeskampPRL88} and then, since the velocities
of electrons at opposite $\mathbf{k}$ points are opposite, a PSC will be
produced. Since the structures presented here are noncentrosymmetric, they are
good candidates in which this effect can be induced.

@@

semiconductors. To derive what we presented above we used

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of
the electron velocity on the plane of the structure,
with the spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\gs^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
to the incoming polarization, 
respectively. The subscript $\gs^\rmb$ denotes the spin along $\rmb$.

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H & -0.61516 & -1.77416 &  0.73196 \\
H &  0.61518 &  0.35514 &  0.73175 \\
C & -0.61516 & -1.77264 & -0.49138 \\
C & -0.61516 & -0.35600 & -0.72316 \\
C &  0.61516 &  0.35763 & -0.49087 \\
\hline
\end{tabular}

\caption{Unit cell of Up structure. Layer division, atom types and
positions for the Up structure. The structure unit cell was divided in
two layers corresponding to hydrogen and carbon atoms.The corresponding layer
atom position is depicted in Fig. \ref{fig:up-struc} with the corresponding
number of layer.}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{cccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H &  -0.61516 &  -1.42140 & \ 1.47237 \\
C &  -0.61516 &  -1.73300 & \ 0.39631 \\
C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
C &  -0.61516 &  -0.37396 &  -0.39632 \\
H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}

@@

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ vs. $\hbar\go$ for the directions $\rma\rmb$, and for the
angle $\ga$
that maximizes the signal, for the Up  and Alt structures and for CdSe and
GaAs, which are bulk systems shown for comparison.
% 
As expected from the delta function of Eq.~\eqref{eq:mu},
$\calv^{\rma\rmb}(\go,\ga)$ rises right at the corresponding energy gap of each
system. For the 2D structures, the spectrum covers two narrow energy regions
with large values of response, while for bulk systems, the spectra covers a
rather wide energy range, but with a much smaller response.
% 
For the Up structure, $\rma\rmb=yz$ and $\ga=35^\circ$ maximize the response,
which means that an incoming beam of light with its electric field polarized at
35$^\circ$ from the $x$ direction will induce electrons to move along $y$
(parallel to the surface), with their spin polarized along $z$ (perpendicular to
the surface), with the following speeds.
% 
Right at the energy onset, $\calv^{yz}(\go,\ga)=668$\,Km/s, a speed that remains
almost constant for 65 meV, and then goes to zero. A second region is found
above 1.946 eV, where there are two extreme values of the speed,
$\calv^{yz}(\go,\ga)=266.3$ Km/s at $\hbar\go=1.954$\,eV, and
$\calv^{\rma\rmb}(\go,\ga)=-241.4$ Km/s at $\hbar\go=1.958$\,eV; a positive
(negative) $\calv^{\rma\rmb}(\go,\ga)$ means that the electrons move parallel
(antiparallel) to the electric field.
Likewise, for the Alt structure, we find that also $\rma\rmb=yz$ and
$\ga=150^\circ$ maximizes the response, where two extreme values of
$\calv^{\rmy\rmz}$ are found, one at  $\hbar\go=0.720$\,eV of
$\calv^{\rmy\rmz}=-711.9$ Km/s, and the other at $\hbar\go=0.911$\,eV of
$\calv^{\rmy\rmz}=-330.6$ Km/s.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{upplots/up-vx-vy-w2}
\caption{For the Up structure we show $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and $\varphi_{\mathrm{a}}
(\omega,\alpha)$, (right scale, blue line), vs. $\hbar\go$, for
$\ga=35^\circ$, and $\rma=x,y$. }
\label{fig:up-vx-vy-w2}
\end{figure}
In Fig. \ref{fig:up-vx-vy-w2}, we plot $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ vs. $\hbar\go$, in the range where there are two local maxima
at $\hbar\go=1.954$\,eV and $\hbar\go=1.957$\,eV.
% 
The former is the largest of the two, with
$\mathcal{V}_{\mathrm{x}} (\omega,\alpha)=61.2$\,Km/s,
$\theta_{\mathrm{x}} (\omega,\alpha)=48.3^{\circ}$, and 
$\varphi_{\mathrm{x}} (\omega,\alpha)=54.3^{\circ}$,
for the electron moving along $x$, and
$\mathcal{V}_{\mathrm{y}} (\omega,\alpha)=293.2$\,Km/s,
$\theta_{\mathrm{y}} (\omega,\alpha)=49.8^{\circ}$, and 
$\varphi_{\mathrm{y}} (\omega,\alpha)=51.9^{\circ}$
for the electron moving along $y$.
% 
For the $\hbar\go=1.957$\,eV peak, we obtain 
$\theta_{\mathrm{x}} (\omega,\alpha) = 129.8^{\circ}$ and 
$\varphi_{\mathrm{x}} (\omega,\alpha) = 231.7^{\circ}$, with 
$\mathcal{V}^{\mathrm{x}} (\omega,\alpha) = 54.6$\,Km/s and
$\theta_{\mathrm{y}}(\omega,\alpha) =129.3$, and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 230.7$, with 
$\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 263.7$\,Km/s.
We remark that these angles are almost
constant for all the energy values across the peak of these two local maxima, for
which the spin is directed upward in the first Cartesian quadrant of the $xy$
plane when it moves along either $x$ or $y$ directions.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{altplots/alt-vx-vy-w1}
\caption{For the Alt structure, we show
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line),
$\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and
$\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue line), vs.
$\hbar\go$, for $\ga=150^\circ$, and $\rma=x,y$. }
\label{fig:alt-vx-vy-w1}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{altplots/alt-vx-vy-w2}
\caption{For the Alt structure, we show $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and $\varphi_{\mathrm{a}}
(\omega,\alpha)$, (right scale, blue line), vs. $\hbar\go$, for
$\ga=150^\circ$, and $\rma=x,y$. }
\label{fig:alt-vx-vy-w2}
\end{figure}

In Figs. \ref{fig:alt-vx-vy-w1} and \ref{fig:alt-vx-vy-w2}, we plot
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line),
$\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and
$\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue line), vs.
$\hbar\go$ in two different ranges, and  for $\rma=x,y$. In this case,
$\ga=150^\circ$  maximizes both $\mathcal{V}_{\mathrm{x}} (\omega,\alpha)$
and $\mathcal{V}_{\mathrm{y}} (\omega,\alpha)$, as a function of $\ga$.
% 
In Fig. \ref{fig:alt-vx-vy-w1}, the absolute maximum is at
$\hbar\go=0.720$\,eV, with $\mathcal{V}_{\mathrm{x}}
(\omega,\alpha) = 301.7$\,Km/s, $\theta_{\mathrm{x}} (\omega,\alpha) =
44.5^{\circ}$ and $\varphi_{\mathrm{x}}(\omega,\alpha) = 51.2^{\circ}$,
and $\mathcal{V}_{\mathrm{y}} (\omega,\alpha) = 905.6$\,Km/s, with
$\theta_{\mathrm{y}} (\omega,\alpha) = 119.7^{\circ}$ and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 163.4^{\circ}$. 
% 
Thus, the spin is directed upward the fourth Cartesian quadrant of the $xy$
plane when the spin velocity is directed along $x$, and directed downward the
second Cartesian quadrant when the spin velocity is directed along $y$.
% 
Finally, in  Fig. \ref{fig:alt-vx-vy-w1}, the absolute maximum is at
$\hbar\go=0.911$\,eV, with $\mathcal{V}_{\mathrm{x}} (\omega,\alpha) =
276.3$\,Km/s, $\theta_{\mathrm{x}} (\omega,\alpha) = 154.6^{\circ}$, and
$\varphi_{\mathrm{x}}(\omega,\alpha) = 292.3^{\circ}$, and
$\mathcal{V}_{\mathrm{y}} (\omega,\alpha) = 468.6$\,Km/s, with
$\theta_{\mathrm{y}} (\omega,\alpha) = 129.2^{\circ}$, and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 228.3^{\circ}$, implying that the spin
is directed downward the fourth Cartesian quadrant of the $xy$ plane when the
spin velocity is directed along $x$, and directed downward the third Cartesian
quadrant when the spin velocity is directed along $y$.
